Short paper : Automatic predicate testing in formal certification

### 2016-04-28 ProofsAndTests/paper-TAP/content/1-introduction.tex
TAP paper

One way to increase the confidence in the software we build is to formally prove its correctness using a proof assistant. Proofs assistants enable to write code, logical statements and proofs in the same language, and offer the guarantee that every proof will be automatically checked. Many of them are functional programming languages, like Coq, Idris and Agda, and others, like the B-Toolkit belong to the imperative paradigm. These different paradigms are internally supported by different logic. Systems like Coq, Idris and Agda are based on various higher order logics (CoC, a variant of ML and LUO respectively) and are realisations of the Curry-Howard correspondence, while the formal B method is based on Hoare logic. These different foundations lead to different philosophies and different ways to implement and verify a software, but all of them greatly increase the confidence of the produced software. However, these guarantees tend to be too often considered as perfect, when they are in fact far from it. Knuth was --certainly ironically-- saying ``Beware of bugs in the above code; I have only proved it correct, not tried it". The reality is precisely that a proof is not enough. When we prove the correctness of a function, we only gain the guarantee expressed by the proven lemma, and nothing more.

@@

We use Idris~\cite{brady2013idris}, a dependently typed programming language, but all the ideas that we present here can be applied to any proof assistant based on type theory.

>>>>

One way to increase the confidence in the software we build is to formally prove its correctness using a proof assistant. Proofs assistants enable to write code, logical statements and proofs in the same language, and offer the guarantee that every proof will be automatically checked. Many of them are functional programming languages, like Coq~\cite{Bertot2004}, Idris~\cite{brady2013idris} and Agda~\cite{Norell2008}, and others, like the B-Method~\cite{Abrial1991} belong to the imperative paradigm. These different paradigms are internally supported by different logic. Systems like Coq, Idris and Agda are based on various higher order logics (CoC, a variant of ML and LUO respectively) and are realisations of the Curry-Howard correspondence, while the formal B method is based on Hoare logic. These different foundations lead to different philosophies and different ways to implement and verify a software, but all of them greatly increase the confidence of the produced software. However, these guarantees tend to be too often considered as perfect, when they are in fact far from it. Knuth was --certainly ironically-- saying ``Beware of bugs in the above code; I have only proved it correct, not tried it". The reality is precisely that a proof is not enough. When we prove the correctness of a function, we only gain the guarantee expressed by the proven lemma, and nothing more.

@@

We use Idris, a dependently typed programming language, but all the ideas that we present here can be applied to any proof assistant based on type theory. The running example that we use in this paper can be found online at \url{https://github.com/FranckS/}.

### 2016-04-28 ProofsAndTests/paper-TAP/content/4-differentKindOfProperties.tex
TAP paper

In order to go a step forward, we can decide to replace the tests on the predicate by proofs. Instead of testing the predicate on a finite subset of all the generated sorted lists, as we did in the previous section, we can prove that any of the automatically generated sorted list is provably sorted according to $isSorted$.

>>>>

In order to go a step forward, we can decide to replace the tests on the predicate by proofs. Instead of testing the predicate on a finite subset of all the generated sorted lists, as we did in the previous section, we can try to prove that any of the automatically generated sorted list is provably sorted according to $isSorted$.
